---
title: "MSDS 6306 - Case Study 1"
author: "Anand Rajan & Kristen McCrary"
date: "10/18/2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

In 2017, 83 percent of all beer was domestically produced, and 17 
percent was imported from more than 100 different countries around 
the world. Based on beer shipment data and U.S. Census population 
statistics, U.S. consumers 21 years and older consumed 26.9 gallons 
of beer and cider per person during 2017. -  
  https://www.nbwa.org/resources/industry-fast-facts

Consumers have many options on the choice of beer they drink today. 
The beer industry is saturated with regards to production and 
consumption over the past 10 years, to ward off competition, and 
position in the market, it is imperative that the brewing companies 
have to come up with new strategies. However, Large breweries still 
are having the majority market share.

Now this study is conducted to help one of our clients on a marketing 
campaign. This analysis would determine certain important questions 
on the Beer and Breweries information for a calculated and appropriate 
decision. The analysis would also include Alcoholic content and IBU 
(international bitterness unit) to position the product with competition 
from other beer manufacturers including craft beer industry. Enough 
sample data is available to conduct the study/analysis.

# Background

Our client Budweiser limited is eager to introduce a new Beer with optimum alcoholic 
content (ABV) and international bitterness units (IBU), so that they can compete 
with other breweries and position itself in the market with regards to pricing. 
This will be an important feature in their marketing campaign and advertisement 
for the product in the upcoming NFL halftime commercial (Most watched) with enhanced 
visual effects.  The strategy would create awareness and curiosity around this new 
beer with their consumers (Budweiser enthusiasts!).

# Analysis

```{r}
# Read in data files
beers <- read.csv("DataFiles/beers.csv")
breweries <- read.csv("DataFiles/breweries.csv")
```

1. How many breweries are present in each state?

```{r}
library(dplyr)
plyr::count(breweries, 'State')
```

2. Merge beer and breweries data

```{r}
beer_data <- merge(breweries, beers, by.x='Brew_ID', by.y='Brewery_id', all=TRUE)
head(beer_data)
tail(beer_data)
```

3. Report number of NA's in each column

```{r}
# Get all the NA data from Beer Data
sapply(beer_data, function(x) sum(is.na(x)))

```
* ABVs - 62   NULLS
* IBUs - 1005 NULLS

# Conclusion

TODO
