---
title: "MSDS 6306 - Case Study 1"
author: "Anand Rajan & Kristen McCrary"
date: "10/18/2018"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
```

# Introduction

"In 2017, 83 percent of all beer was domestically produced, and 17 percent was imported from more than 100 different countries around the world. Based on beer shipment data and U.S. Census population statistics, U.S. consumers 21 years and older consumed 26.9 gallons of beer and cider per person during 2017." (Source: https://www.nbwa.org/resources/industry-fast-facts)

Consumers have many options on the choice of beer they drink today. The beer industry is saturated with regards to production and consumption over the past 10 years. To ward off competition, and to earn a position in the market, it is imperative that brewing companies come up with new strategies. However, large breweries still have the majority market share.

This study will help one of our clients on a marketing campaign. This analysis will determine certain important questions on the Beer and Breweries information for a calculated and appropriate decision. The analysis will also include alcoholic content and IBU (international bitterness unit) to position the product against competition from other beer manufacturers, including the craft beer industry. Enough sample data is available to conduct the study/analysis.

# Background

Our client Anheuser Busch is one of the top beer producers with the following vitals to boast. 
Head quartered in St Louis MO, the company brews more than 100 brands of beers with flag ship brands Budweiser and Budlight.
The company also owns 23 breweries and has several craft partners to its credit. The company's revenue stands at 14 Billion dollars per year.
Currently the company is eager to introduce a new beer with optimum alcoholic content (ABV) and international bitterness units (IBU), so that they can compete with other breweries and position itself in the market with regards to pricing. 

# Analysis

To aid in our analysis, our client provided us with two highly relevant datasets. One dataset contains a list of 2410 craft beers brewed in the U.S., while the other contains information on 558 competing U.S. breweries. As we are using the programming language R to perform our analysis, we first have to load these datasets into R. In order to show some good analysis, we had cleaned up the data to show information where the data is available.

```{r}
# Read in beer and brewery data files
beers <- read.csv("DataFiles/beers.csv")
breweries <- read.csv("DataFiles/breweries.csv")
```

#### 1. State-wise brewery counts

Our client would first like to know how many breweries are present in each state. This will give the company an idea of where their competition is geographically located.

```{r fig.align="center"}
# Count number of breweries in each state
cdf <- plyr::count(breweries, 'State')
names(cdf)<- c("State", "NumBreweries")

# Define color gradient
grad <- scales::seq_gradient_pal("brown", "yellow")(seq(0,1,length.out=51))

# Create barchart for breweries per state
ggplot(cdf, aes(x=reorder(State, -NumBreweries), y=NumBreweries, fill=State)) +
  geom_bar(stat='identity', position='dodge') +
  labs(title="Figure 1: Number of Breweries per State", x="State", y="Number of Breweries") +
  theme(plot.title = element_text(hjust=0.5), axis.text.x=element_text(angle=90, size=7), 
        legend.position="none") +     
  scale_fill_manual(values=grad)
```

The Figure 1 barchart shows the number of breweries categorized by states in the U.S. (plus District of Columbia). We see that Colorado leads with 47 breweries, followed by California and Michigan. 

#### 2. Merge beer and breweries data

To fully take advantage of the two datasets, we merge them into one composite dataset. They can be combined because each beer is brewed at, or at least associated with, a particular brewery.

```{r}
# Merge data on brewery ID fields
beer_data <- merge(breweries, beers, by.x='Brew_ID', by.y='Brewery_id', all=TRUE)
# Rename ambiguous columns
names(beer_data)[c(2, 5)] <- c("Brewery_Name", "Beer_Name")

# Display beginning of merged data frame
# Hide ID fields, not necessary for client to see
hd <- select(head(beer_data), -Brew_ID, -Beer_ID)
kable(hd, row.names=FALSE, caption="Table 1: Beginning of Merged Data Frame") %>% kable_styling(bootstrap_options="bordered")

# Display end of merged data frame 
# Hide ID fields, not necessary for client to see
tl <- select(tail(beer_data), -Brew_ID, -Beer_ID) # Hide ID fields
kable(tl, row.names=FALSE, caption="Table 2: End of Merged Data Frame") %>% kable_styling(bootstrap_options="bordered")
```

Table 1 and Table 2 show the first and last six observations of the combined file, repectively, showing that the merge was a success. We notice that breweries are often repeated in this new dataset, since breweries often make multiple beers.

#### 3. Report column NA's

For our analysis, we would also like to know the number of NA's, i.e. missing values, in each column. In this way we can be aware that some inaccuracies might occur because not all of the data on each beer and brewery is available.

```{r fig.align="center"}
# Get all the NAs from beer data
na_table <- sapply(beer_data, function(x) sum(is.na(x)))
na_df <- as.data.frame(na_table)

# Create bar chart of missing values
ggplot(na_df, aes(x=reorder(rownames(na_df), na_table), y=na_table)) + 
  geom_bar(stat='identity', position='dodge', fill="darkred", color="darkred") +
  coord_flip() +
  labs(title="Figure 2: Missing Values", x="Variable", y="Number of NAs") +
  theme(plot.title = element_text(hjust=0.5), legend.position="none") 

# Create "kable" of values
kbl <- kable(na_table, col.names="Number of NAs", caption="Table 3: NA Counts")
kable_styling(kbl, full_width=FALSE, bootstrap_options="bordered")
```

From Figure 2 and Table 3 we see that there are 62 NA values in the ABV column, and 1005 NA's in the IBU column. There are no null values for the remaining variables. 

#### 4. Median ABV and IBU by state

We would now like to visualize the median alcohol content and bitterness for beers in each state. 

```{r fig.align="center"}
# Get median values for each state, convert to data frames
median_ABV <- tapply(beer_data$ABV, beer_data$State, median, na.rm=TRUE)
median_IBU <- tapply(beer_data$IBU, beer_data$State, median, na.rm=TRUE)
median_ABV <- as.data.frame(median_ABV); median_ABV$State <- rownames(median_ABV)
median_IBU <- as.data.frame(median_IBU); median_IBU$State <- rownames(median_IBU)
# Remove South Dakota since no IBU data was available
median_IBU <- subset(median_IBU, State!=" SD")

# Define color gradient
grad <- scales::seq_gradient_pal("brown", "yellow")(seq(0,1,length.out=51))

# Create barchart for median ABV
ggplot(median_ABV, aes(x=reorder(State, -median_ABV), y=median_ABV, fill=State)) +
  geom_bar(stat='identity', position='dodge') +
  labs(title="Figure 3: Median Alcohol Content of Beers by State", x="State", y="Median ABV") +
  theme(plot.title = element_text(hjust=0.5), axis.text.x=element_text(angle=90, size=7), 
        legend.position="none") +     
  scale_fill_manual(values=grad)

# Create barchart for median IBU
ggplot(median_IBU, aes(x=reorder(State, -median_IBU), y=median_IBU, fill=State)) +
  geom_bar(stat='identity', position='dodge', na.rm=TRUE) +
  labs(title="Figure 4: Median Bitterness of Beers by State", x="State", y="Median IBU") +
  theme(plot.title = element_text(hjust=0.5), axis.text.x=element_text(angle=90, size=7), legend.position="none") +     
  scale_fill_manual(values=grad)
```

The alcohol content in beers by consuming states are plotted in a geometric bar graph, capturing the highest to lowest alcohol by volume and bitterness index. We have two different plots, Figure 3 showing the median ABV and Figure 4 visualizing the median IBU. Note that South Dakota is missing from Figure 4, because there was no data available on bitterness of beers in that state.

#### 5. Maximum ABV and IBU

Next, we want to know the state that has the beer with the largest alcohol content, as well as the state containing the most bitter beer. 

```{r}
# Get top 3 beers for alcohol content
max_ABV <- beer_data[order(beer_data$ABV, decreasing=TRUE), c("State", "Beer_Name", "ABV", "Brewery_Name")] %>% head(n=3)
kbl <- kable(max_ABV, row.names=FALSE, caption="Table 4: Highest ABV")
kable_styling(kbl, full_width=FALSE, bootstrap_options="bordered")

# Get top 3 most bitter beers
max_IBU <- beer_data[order(beer_data$IBU, decreasing=TRUE), c("State", "Beer_Name", "IBU", "Brewery_Name")] %>% head(n=3)
kbl <- kable(max_IBU, row.names=FALSE, caption="Table 5: Highest IBU")
kable_styling(kbl, full_width=FALSE, bootstrap_options="bordered")
```

As seen from the Table 4, the state of Colorado has the maximum alcohol content for beers at 12.8%. In addition, from Table 5 we see that Oregon has the most bitter beer at 138 bitterness units.

#### 6. ABV summary statistics

We will see the summary statistics of alcohol by volume for all beers in the United States.

```{r fig.align="center"}
# Get summary statistics
sum <- as.array(summary(beer_data$ABV)) %>% round(3)

fun_mean <- function(x){
  return(data.frame(y=mean(x),label=mean(x,na.rm=T)))}

# Make boxplot, including mean
ggplot(beer_data, aes(y=ABV, x="")) + 
  geom_boxplot(na.rm=TRUE, width=0.4, color="#31394d") +
  stat_summary(fun.y=mean , na.rm=TRUE, geom="point", colour="darkred") + 
  coord_flip() + 
  labs(title="Figure 5: Alcohol Content Summary", x="", y="Alcohol by Volume (ABV)") +
  theme(plot.title = element_text(hjust=0.5))

# Make table of values
kbl <- kable(sum, col.names=c("Statistic", "ABV Value"), caption="Table 6: ABV Summary Statistics") 
kable_styling(kbl, full_width=FALSE, bootstrap_options="bordered") %>% column_spec(1:2, width="2.75cm")
```

The boxplot in Figure 5 and summary in Table 6 includes the minimum, median, mean, and maximum of alcohol by volume (ABV). In percentage of alcohol content these values represent 0.1%, 5.6% , 6% and 12.8% by volume, respectively.

#### 7. IBU and ABV relationship

```{r fig.align="center"}
# Create scatterplot
ggplot(beer_data, aes(x=IBU, y=ABV, color=IBU)) + 
  geom_point(size=1.3, na.rm=TRUE) + 
  geom_smooth(method=lm, na.rm=TRUE, se=FALSE, color="brown") +
  labs(title="Figure 6: Alcohol Content vs. Bitterness of Beers", x="International Bitterness Units (IBU)", y="Alcohol by Volume (ABV)") +
  theme(plot.title = element_text(hjust=0.5), legend.position="none") +
  scale_color_gradient(low = "#ffbf00", high = "brown")

# Get correlation coefficient
with(beer_data, cor.test(IBU, ABV))
```

The scatter plot in Figure 6 shows a moderately strong positive linear correlation between the IBU and ABV of beers, which is supported by the correlation coefficient of 0.67. As bitterness increases we see that the alcohol content generally increases as well, barring some outliers. The line drawn through the scatter plot represents the line of best fit, which enables us to predict alcohol content from a given bitterness level based on the data.

# Conclusion

Considering the given information and the market trends thus analyzed so far, the conclusion is the CEO can take a decision on launching a new product in collaboration with the craft beer breweries. The new product should have optimum ABV and IBU to satisfy the current consumer trends in beer drinking as they prefer like light lagers and craft products.
